---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
## Set the global option echo to true, so that the code is always visible.
```{r setoptions, echo = TRUE}
library(knitr)
opts_chunk$set (echo = TRUE, results = TRUE)
```

## Download the data into the workspace in case it's not already there.
```{r}
download.file('https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip', 'activity.zip')
```

## Loading and preprocessing the data
```{r}
unzip('activity.zip')
data <- read.csv2('activity.csv', sep = ',')
```

## What is mean total number of steps taken per day?
### First subset out all the observations with NA steps.
```{r}
data2 <- na.omit(data)
```

### Secondly calculate the total steps for each day.
```{r}
library(dplyr)
df <- data2 %>% group_by(date) %>% summarise_each(funs(sum), steps)
df_total_steps <- sum(df$steps)
```

### Thirdly, make a historgram of the total number of steps per day.
```{r}
hist(df$steps)
```

### Finally, calculate the mean and median of the total num of steps taking per day.

``` {r}
df_mean <- mean(df$steps)

df_median <- median(df$steps)

df_mean
df_median
```
The mean total number of steps taken per day is `r df_mean` and the median of the total number of steps taken per day is `r df_median`.

## What is the average daily activity pattern?
### First change the intervals into factors.
```{r}
data3 <- data2
data3$interval <- as.factor(data3$interval)
```

### Secondly, find the average number of steps taken, averaged across all days.
```{r}
df2 <- data3 %>% group_by(interval) %>% summarise_each(funs(mean), steps)
```

### Finally, plot the resulting data.
```{r}
plot(df2$interval, df2$steps, type = '1')
```
### Find the 5 minute interval with the maximum number of steps.
```{r}
max_interval <- df2[which.max(df2$steps),1]
```
The `r max_interval` 5-minute interval, on average, contains the maximum number of steps.

## Imputing missing values
```{r}
summary(data)
```

The total number of missing values in the dataset is 2304.

The startegy we will use to fill in the missing values is to use the mean for that specific day.
We will use the Hmisc package's impute function.

``` {r}
library(Hmisc)
data4 <- data
data4$steps <- impute(data4$steps)
df3 <- data4 %>% group_by(date) %>% summarise_each(funs(sum), steps)
hist(df3$steps)

df3_mean <- mean(df3$steps)

df3_median <- median(df3$steps)

df3_mean
df3_median
```
The adjusted mean total number of steps taken per day is `r df3_mean` and the adjusted median of the total number of steps taken per day is `r df3_median`.  The previous calculations were `r df_mean` and `r df_median`.

``` {r}
df3_total_steps <- sum(df3$steps)
```

The adjusted total number of steps is `r df3_total_steps` while the previous number was `r df_total_steps`.


## Are there differences in activity patterns between weekdays and weekends?
### Convert the dates into weekday factors.
Note that I am only considering Saturday and Sunday as weekends. THe rest are weekdays.
```{r}
data5 <- data4
weekdays <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
data5$date = as.factor(ifelse(is.element(weekdays(as.Date(data5$date)),weekdays), "Weekday", "Weekend"))
```

Calculate the average of each interval over all weekdays and weekends.


###
```{r}
library(ggplot2)
df_final <- data5 %>% group_by(interval, date) %>% summarise_each(funs(mean), steps)
ggplot(df_final, aes(interval, steps)) + geom_line() + facet_grid(~ date)

```